---
title: "lab02-visualizations"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: true
    allow_skip: true
    toc: true
    toc_depth: 3
    theme: readable
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
# library(magrittr)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

## Why should we visualize networks?
The visualization of network is a specific field of data visualization. In general
we visualize data for several reasons, but one of the most important is to make 
data and results clear and understandable for a larger public. Communicability 
is a very important skill both in academia and in the private sector. 

You might have very important findings, but if your audience does not understand
them they are worth nothing. 
A great visualization makes more than half of the job! 
Great visualizions are story tellers!

### Goals of network visualization

* to identify key actors and the relationship between them
* to spot most influential actors
* to check the strength of relationships
* to see whether people belong to sub-groups inside large networks 
* to analyze the structural properties of networks
* to unfold diffusion patterns
* to observe network evolution
* to summaries any sort of relationship among entities in one picture
* ... and many more reasons!


## Basic network plots

Plotting networks is essentially plotting dots and lines. The R environment allow
us to do it already.
```{r}
base::plot(x=1:5, y=rep(5,5), pch=19, cex=12, col=rgb(.25, .5, .3, alpha=.5), xlim=c(0,6)) 
```
However, it is very difficult to do it in the base R programming language. Hence 
there are a few packages that help us on the job. 

### Basic plots
As much as for the majority of network analysis task, we can use two packages

* igraph
* networks


Let's check a first example with igraph
```{r}
library(igraph)
set.seed(10000)
igraphNet <- igraph::erdos.renyi.game(10, 2/10)
plot(igraphNet)
```

this is the most basic network plot we can have. 

We can do a very similar one using network 
```{r}
detach("package:igraph", unload = TRUE)
library(intergraph)
networkNet <- intergraph::asNetwork(igraphNet)
plot(networkNet)
```
Different layout set up, same output (fortunately... otherwise it would mean we
made some mistakes somewhere)


However, for more complex networks to plot, we need to make our visualization
skills more sophisticated
```{r}
library(igraph)
set.seed(10001)
igraphNetLarge <- igraph::erdos.renyi.game(1000, 1/1000)
plot(igraphNetLarge)

```


this last plot does not meet any of the goal we stated above. It's just a mess!


```{r testquiz}
question("What is the main problem with the plots of igraphNet and networkNet?",
    answer("nodes are too small"),
    answer("the colors are not color blind friendly"),
    answer("the plot does not inform us on what the network represents", correct = TRUE),
  )
```

yes, even if these two example plots are readable, they are not very helpful in 
telling a story!

## Tell your network story

In this first part of the tutorial we will explore the options for plotting
networks in igraph.

We will use one of the data sets from a tutorial that I kindly 
recommend: 

Ognyanova, K. (2019) Network visualization with R. 
Retrieved from www.kateto.net/network-visualization.

It's a media network
```{r}
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)

detach("package:intergraph", unload = TRUE)
library(igraph)
net <- graph_from_data_frame(d=links, vertices=nodes, directed=T) 
```
Loading the network from a data frame allows the automatic upload of several 
information about the nodes and their relationships.

Let's briefly explore the network
```{r}
vcount(net)
```
What do nodes represent? We need to know the name of the attribute we want to
display, and we call it using the $ sign.
```{r}
V(net)$media
```
How many relationships in total we observed between these media?
```{r}
ecount(net)
```

For a preliminary overview this is enough. We already have a clue of what we are 
dealing with. 

Let's plot!
```{r}
plot(net) 
```
we can do better than this... let's remove loops and plot it reducing the arrow 
size, removing the vertex label and curving the edges.
```{r}
net <- igraph::simplify(net, remove.multiple = F, remove.loops = T) 
plot(net, 
     edge.arrow.size=.4, 
     vertex.label=NA, 
     edge.curved=.1)
```
it's already prettier, but still not very informative. 

Printing the media attribute instead of the label makes clearer what we are 
trying to communicate. Let's do it, together with some aesthetic improvements!
```{r}
plot(net, 
     edge.arrow.size=.2, 
     edge.color="orange",
     vertex.color="orange", 
     vertex.frame.color="#ffffff",#node perimeter
     vertex.label=V(net)$media, 
     vertex.label.color="black") 
```
this is a much better job! 

We have more information on this network, let's try to show it. We know the 
audience size for each of those media. We can plot it as a node size! We can 
also change color according to media type. 
```{r}
# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type] # we attribute media type to color

V(net)$size <- V(net)$audience.size*0.6 # and audience size to size

# The labels are currently node IDs.
# Setting them to NA will render no labels:
V(net)$label <- NA

#change arrow size and edge color (in the previous example they were set inside 
#the plot function. Both options work)
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"

plot(net,
     vertex.frame.color="#ffffff",#node perimeter
     vertex.label=V(net)$media, 
     vertex.label.color="black"
     ) 
legend(x=-1.5, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```
This plot is very informative! in one glance we can easily see that TV channels 
have a much larger audience than newspapers and online news. 
That the most popular newspaper is the New York Post, and the most 
popular TV channel Fox news. 

It is hard to read the direction of the edges though! We can fix it!
Let’s color the edges of the graph based on their source node color.
```{r}
# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type] # we attribute media type to color

V(net)$size <- V(net)$audience.size*0.6 # and audience size to size

# The labels are currently node IDs.
# Setting them to NA will render no labels:
V(net)$label <- NA

# color the edge as the sender 
edge.start <- ends(net, es=E(net), names=F)[,1]
edge.col <- V(net)$color[edge.start]

plot(net,
     vertex.frame.color="#ffffff",#node perimeter
     vertex.label=V(net)$media, 
     vertex.label.color="black",
     edge.color=edge.col, 
     edge.curved=.1
     ) 
legend(x=-1.5, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)

```


It's interesting to note that BBC is the only non-American media, but it is 
mentioned or linked by the other news source. Weirdly BBC never mentioned anyone,
or more likely, it wasn't part of the analyzed sample, but only cited by the 
American media monitored in this study! 

That's already a large portion of the story these data can tell! 

### Layout
Network layouts are simply algorithms that return coordinates for each node in 
a network.
We can improve the look of our visualization by setting some layout options that
we believe work best with our data.
```{r}
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type] 
V(net)$size <- V(net)$audience.size*0.6 
V(net)$label <- NA
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"

graph_attr(net, "layout") <- layout_with_lgl # layout option added to the same code
plot(net,
     vertex.frame.color="#ffffff",#node perimeter
     vertex.label=V(net)$media, 
     vertex.label.color="black"
     ) 
legend(x=-1.5, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```
it looks much tidier than before! There are several layout options. 
Let's explore a few more

```{r}
colrs <- c("gray50", "tomato", "gold")
V(net)$color <- colrs[V(net)$media.type] 
V(net)$size <- V(net)$audience.size*0.6 
V(net)$label <- NA
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"

graph_attr(net, "layout") <- layout_in_circle # layout option added to the same code
plot(net,
     vertex.frame.color="#ffffff",#node perimeter
     vertex.label=V(net)$media, 
     vertex.label.color="black"
     ) 
legend(x=-1.5, y=-1.1, c("Newspaper","Television", "Online News"), pch=21,
       col="#777777", pt.bg=colrs, pt.cex=2, cex=.8, bty="n", ncol=1)
```

You can play with the code substituting some other layout
* layout_randomly
* layout_on_sphere
* layout_with_fr

Look up these algorithms in the help files to understand them!
There are many more, and we will encounter them during the following weeks. 



## ggraph

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
