---
title: "lab02-visualizations"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: true
    allow_skip: true
    toc: true
    toc_depth: 3
    theme: readable
runtime: shiny_prerendered
    
---

```{r setup, include = FALSE}
library(learnr)
library(gradethis)
# library(magrittr)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

## Why should we visualize networks?
The visualization of networks is a specific field of data visualization.
In general, we visualize data for several reasons, but one of the most 
important is to make data and results clear and understandable for a 
larger public. Communicability is an essential skill both in academia 
and in the private sector. 

You might have very important findings, but if your audience does not
understand them, they are worth nothing. 
A great visualization makes more than half of the job! 
Great visualizations are story tellers!

### Goals of network visualization

* to identify key actors and the relationship between them
* to spot most influential actors
* to check the strength of relationships
* to see whether people belong to sub-groups inside large networks 
* to analyze the structural properties of networks
* to unfold diffusion patterns
* to observe network evolution
* to summaries any relationship among entities in one picture
* ... and many more reasons!


## Basic network plots

Plotting networks is essentially plotting dots and lines. 
The R environment allows us to do it already.
```{r nonNetPlot}
x <- 1:10
y1 <- x*x
y2  <- 2*y1
base::plot(x, y1, type = "b", pch = 19, 
            col = "red", xlab = "x", ylab = "y")
```

However, it isn't easy to represent what you want in the 
R base programming language only.
R base does not recognize a network and does not provide 
a ready-made environment to plot them.

Hence there are a few packages that help us on the job. 

### Basic plots
There are two very popular packages to plot networks

* `igraph`
* `networks`

you met them already

Let's check a first example with `igraph`
```{r randomEs1}
library(igraph)
set.seed(10000)
igraphNet <- igraph::erdos.renyi.game(10, 2/10)
base::plot(igraphNet)
```

This is the most basic network plot we can have. 

We can do a very similar one using `network` 
```{r randomEs2}
detach("package:igraph", unload = TRUE)
library(intergraph)
networkNet <- intergraph::asNetwork(igraphNet)
base::plot(networkNet)
```


Different layout set up, same output (fortunately...
otherwise it would mean we made some mistakes somewhere)


For more complex networks to plot, we need to make our visualization
skills more sophisticated
```{r randomEs3}
library(igraph)
set.seed(10001)
igraphNetLarge <- igraph::erdos.renyi.game(1000, 1/1000)
base::plot(igraphNetLarge)

```


This last plot does not meet any of the goal we stated above. It's just a mess!


```{r testquiz, echo = FALSE}
question("What is the main problem with the plots of igraphNet and networkNet?",
    answer("nodes are too small"),
    answer("the colors are not color blind friendly"),
    answer("the plot does not inform us on what the network represents", correct = TRUE)
  )
```

Yes, even if these two example plots are readable, 
they are not very helpful to tell a story!

## Tell your network story

In this first part of the tutorial, we will explore the options for plotting
networks in `igraph`. The package `network` has very similar tools, but it 
provides a smaller number of customizable options. 

We will use one of the data sets from a tutorial that I kindly 
recommend: 

Ognyanova, K. (2019) Network visualization with R. 
Retrieved from www.kateto.net/network-visualization.

It's a media network
```{r dataload, results = "hide"}
nodes <- utils::read.csv("Dataset1-Media-Example-NODES.csv", header = T, as.is = T)
links <- utils::read.csv("Dataset1-Media-Example-EDGES.csv", header = T, as.is = T)

library(igraph)
net <- igraph::graph_from_data_frame(d=links, vertices=nodes, directed = T) 
```


Loading the network from a data frame allows the automatic upload of large 
quantities of information about the nodes and their relationships.

Let's briefly explore the network.
```{r datacheck}
igraph::vcount(net)
```

What do nodes represent? We need to know the name of the attribute 
we want to display, and we call it using the $ sign.

```{r datacheck1}
igraph::V(net)$media
```

How many relationships in total we observed between these media?
```{r datacheck2}
igraph::ecount(net)
```

For a preliminary overview this is enough. We already have a clue of what we are 
dealing with. 

Let's plot!
```{r plot1}
base::plot(net) 
```

We can do better than this... let's remove loops and plot it reducing the arrow 
size, removing the vertex label and curving the edges.

```{r plot2}
net <- igraph::simplify(net, remove.multiple = F, remove.loops = T) 
base::plot(net, 
           edge.arrow.size = .4, 
           vertex.label = NA, 
           edge.curved = .1)
```


It's already prettier, but still not very informative. 

Printing the media name (one of the attributes) instead of the label 
clarifies what we are trying to communicate. Let's do it, together
with some aesthetic improvements!
```{r plot3}
base::plot(net, 
           edge.arrow.size = .2, 
           edge.color = "orange",
           vertex.color = "orange", 
           vertex.frame.color = "#ffffff", # node perimeter
           vertex.label = igraph::V(net)$media, 
           vertex.label.cex = 0.6,  # vertex label size
           vertex.label.color = "black") 
```

This is a much better job! 

We have more information on this network, let's try to show it.
We know the audience size for each of those media. We can plot 
it as a node size! We can also change color according to media type. 
```{r plot4}
# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
igraph::V(net)$color <- colrs[igraph::V(net)$media.type] # we attribute media type to color

igraph::V(net)$size <- igraph::V(net)$audience.size*0.6 # and audience size to size

# The labels are currently node IDs.
# Setting them to NA will render no labels:
igraph::V(net)$label <- NA

#change arrow size and edge color (in the previous example they were set inside 
#the plot function. Both options work)
igraph::E(net)$arrow.size <- .2
igraph::E(net)$edge.color <- "gray80"

base::plot(net,
           vertex.frame.color = "#ffffff",#node perimeter
           vertex.label = igraph::V(net)$media, 
           vertex.label.cex = 0.6,  # vertex label size
           vertex.label.color = "black"
) 
graphics::legend(x = -1.5, y = -1.1, c("Newspaper","Television", "Online News"), pch = 21,
                 col = "#777777", pt.bg = colrs, pt.cex = 2, cex = .8, bty = "n", ncol = 1)
```


This plot is very informative! In one glance, we can easily see that TV 
channels have a much larger audience than newspapers and online news. 
That the most popular newspaper is the New York Post, and the most 
popular TV channel Fox news. 

It is hard to read the direction of the edges, though! We can fix it!
Let’s color the edges of the graph based on their source node color.
```{r plot5}
# Generate colors based on media type:
colrs <- c("gray50", "tomato", "gold")
igraph::V(net)$color <- colrs[igraph::V(net)$media.type] # we attribute media type to color

igraph::V(net)$size <- igraph::V(net)$audience.size*0.6 # and audience size to size

# The labels are currently node IDs.
# Setting them to NA will render no labels:
igraph::V(net)$label <- NA

# color the edge as the sender 
edge.start <- igraph::ends(net, es = igraph::E(net), names = F)[,1]
edge.col <- igraph::V(net)$color[edge.start]

base::plot(net,
           vertex.frame.color = "#ffffff",#node perimeter
           vertex.label = igraph::V(net)$media, 
           vertex.label.cex = 0.6,  # vertex label size
           vertex.label.color = "black",
           edge.color = edge.col, 
           edge.curved = .1
) 
graphics::legend(x = -1.5, y = -1.1, c("Newspaper","Television", "Online News"), pch = 21,
                 col = "#777777", pt.bg = colrs, pt.cex = 2, cex = .8, bty = "n", ncol = 1)

```


It's interesting to note that the BBC is the only non-American media, 
but it is mentioned or linked by other news sources. Weirdly BBC never 
mentioned anyone, or more likely, it wasn't part of the analyzed sample, 
but only cited by the American media monitored in this study! 

That's already a large portion of the story these data can tell! 

### Layout

Network layouts are simply algorithms that return coordinates 
for each node in a network.
We can improve the look of our visualization by setting some 
layout options that we believe work best with our data.

```{r plot6}
colrs <- c("gray50", "tomato", "gold")
igraph::V(net)$color <- colrs[igraph::V(net)$media.type] 
igraph::V(net)$size <- igraph::V(net)$audience.size*0.6 
igraph::V(net)$label <- NA
igraph::E(net)$arrow.size <- .2
igraph::E(net)$edge.color <- "gray80"

graph_attr(net, "layout") <- layout_with_lgl # layout option added to the same code
base::plot(net,
           vertex.frame.color = "#ffffff",#node perimeter
           vertex.label = igraph::V(net)$media, 
           vertex.label.cex = 0.6,  # vertex label size
           vertex.label.color = "black"
) 
graphics::legend(x = -1.5, y = -1.1, c("Newspaper","Television", "Online News"), pch = 21,
                 col = "#777777", pt.bg = colrs, pt.cex = 2, cex = .8, bty = "n", ncol = 1)
```


It looks much tidier than before! There are several layout options. 
Let's explore a few more

```{r plot7}
colrs <- c("gray50", "tomato", "gold")
igraph::V(net)$color <- colrs[igraph::V(net)$media.type] 
igraph::V(net)$size <- igraph::V(net)$audience.size*0.6 
igraph::V(net)$label <- NA
igraph::E(net)$arrow.size <- .2
igraph::E(net)$edge.color <- "gray80"

graph_attr(net, "layout") <- layout_in_circle # layout option added to the same code
base::plot(net,
           vertex.frame.color = "#ffffff",#node perimeter
           vertex.label = igraph::V(net)$media, 
           vertex.label.cex = 0.6,  # vertex label size
           vertex.label.color = "black"
) 
graphics::legend(x = -1.5, y = -1.1, c("Newspaper","Television", "Online News"), 
                 pch = 21, col = "#777777", pt.bg = colrs, pt.cex = 2, cex = .8, bty = "n", ncol = 1)
```

You can play with the code substituting some other layout
* layout_randomly
* layout_on_sphere
* layout_with_fr

Look up these algorithms in the help files to understand them!
There are many more, and we will encounter them during the following weeks. 



## ggraph

`ggplot2` is the most popular package for visualization in R but it does not 
handle network data. The `ggraph` package address the problem expanding 
`ggplot2`. 

The logic behind `ggplot2` consists of building plots adding features to 
a structure layer by layer. 


### A bit with ggplot2

Let's see an example with the toy data set 'mtcars' from the 
package `datasets`. 

Shall we take a look?

```{r dataNonNet}
utils::data(mtcars, package = "datasets")
head(mtcars) # let's have a look to the data
```


First we create a canvas
```{r ggplotexample}
library(ggplot2)
ggplot2::ggplot(mtcars)
```


Then we had the variable of interest

```{r ggplotexample1}
# variables of interest mapped
ggplot2::ggplot(mtcars, ggplot2::aes(x = disp, y = wt))
```

Now we plot the data adding geometric points
```{r ggplotexample2}
ggplot2::ggplot(mtcars, ggplot2::aes(x = disp, y = wt)) +
  ggplot2::geom_point()
```

Layer by layer we can make it as pretty as we like
```{r ggplotexample3, message = FALSE, warning = FALSE}
ggplot2::ggplot(mtcars, ggplot2::aes(x = disp, y = wt)) +
  ggplot2::geom_point(ggplot2::aes(color = gear)) + # the colour expresses one chosen attribute
  ggplot2::geom_smooth(se = TRUE) +  # adding the line that minimizes the distance and
  # display the standard error
  ggplot2::theme_bw()+ # changing the background colour
  ggplot2::ggtitle("Some mtcars plot I like")

  
```

It's (almost) completely customizable


### Let's try ggraph

We are lucky, since `ggraph` reads `igraph` format. We can just keep plotting 
our media network. Let's create a canvas first.
```{r ggpraphexample, message = FALSE, warning = FALSE}
library(ggraph)
ggraph::ggraph(net) 
```

Let's add nodes and adges now

```{r ggpraphexample1, message = FALSE, warning = FALSE}
ggraph::ggraph(net) +
  ggraph::geom_edge_link() +   # add edges to the plot
  ggraph::geom_node_point()    # add nodes to the plot
```

This is already pretty decent I would say! A specific layout is already set.
But we can easily change it!
```{r ggpraphexample2, message = FALSE, warning = FALSE}
ggraph::ggraph(net, layout = "kk") + # changed layout
  ggraph::geom_edge_link() +
  ggraph::geom_node_point() +
  ggplot2::ggtitle("Pretty Plot in Progress") + # same as in ggplot2!
  ggplot2::theme_bw() # same as in ggplot2!
```

Let's try something else
```{r ggpraphexample3, echo = FALSE}
ggraph::ggraph(net, layout = "lgl") + # changed layout again
  ggraph::geom_edge_fan(color = "gray50", width = 0.8, alpha = 0.5) + # set the type of edge 
  # to make sure any overlapping multiplex edges will be fanned out.
  ggraph::geom_node_point(color = igraph::V(net)$color, size = 8) + # used the color we set before
  ggplot2::ggtitle("Even Prettier Plot") +
  ggplot2::theme_void() # removed the grid that is not always helpful in network viz
```


We can also set the edges to a curvy shape
```{r ggpraphexample4}
ggraph::ggraph(net, layout = 'linear') + 
  ggraph::geom_edge_arc(color = "orange", width = 0.7) +
  ggraph::geom_node_point(size = 5, color = "gray50") +
  ggplot2::ggtitle("Legit Weird Plot") +
  ggplot2::theme_void()
```

Funny look together with this linear layout right?
But we can be serious too
```{r ggpraphexample5}
library(ggplot2)
ggraph::ggraph(net, layout = "lgl") +
  ggraph::geom_edge_link(aes(color = type)) +           # colors by edge type 
  ggraph::geom_node_point(aes(size = audience.size)) +  # size by audience size  
  ggplot2::theme_void() +
  ggplot2::ggtitle("More informative Plot")
```


Here we can observe the size of the audience from the node size 
and the type of edge from the color. 

We can also observe that a legend is printed automatically. 
However, we can still switch it off if for some reasons we don't want it!
```{r ggpraphexample6, message = FALSE, warning = FALSE}
ggraph::ggraph(net,  layout = 'lgl') +
  ggraph::geom_edge_arc(color = "yellow", curvature = 0.3) +  # we can decide on the curvature  
  ggraph::geom_node_point(color = "blue", aes(size = audience.size)) +     
  ggraph::geom_node_text(aes(label = media), size = 3, color = "gray50", repel = T) +
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none") + # eliminate legend
  ggplot2::ggtitle("Plot in yellow and blue")

```


We can also customize more features

```{r ggpraphexample7}
ggraph::ggraph(net,  layout = 'lgl') +
  ggraph::geom_edge_fan(color = "blue") + # change edge option
  ggraph::geom_node_point(color = "red", aes(size = audience.size)) +     
  ggraph::geom_node_text(aes(label = media), size = 4, color = "black", repel = T) +
  ggplot2::theme_void() +
  ggplot2::theme(legend.key.size = unit(1.5, "cm"),
                 legend.key.width = unit(0.5,"cm"),
                 legend.position = "left",
                 legend.background = element_rect(color = "red", fill = "blue"),
                 legend.title = element_text(family = "arial", color = "white")
  ) + 
  ggplot2::ggtitle("Very customized Plot", subtitle = "Do you like it?") 
  
  
```
Yes, very customizable, but please never plot something so ugly :)



Plotting in `ggraph` is much more flexible than `igraph` and `networks`, since
we can benefit from all the options that are already set for `ggplot2`.
```{r removePackage, echo = FALSE, include = FALSE}
detach("package:ggraph", unload = TRUE)
```


## Dynamic Plots

Telling network stories can be very difficult. Sometimes, we need 
to display how certain networks change over time or show several 
visualizations that provide different information about the same network. 

In these cases, a dynamic plot can help. 

Dynamics plots are essentially the automatic display of more
than one static plot.
We already know how to make static plots; we just need to 
learn how to put a few of them together. 


### Embed dynamic plots on Rmarkdown

If we simply need to embed a dynamic plot on the RMarkdown worksheet 
to show it in html, we have have an easy game to play!

Let's first create a plot and let's store it into a variable
```{r dynamic}
p1 <- ggraph::ggraph(net, layout = 'kk') +
      ggraph::geom_edge_arc(color = "orange", width = 0.7) +
      ggraph::geom_node_point(size = 5, color = "gray50")
p1
```

Let's make the same plot in two more versions using different colours and layout

```{r dynamic1}
p2 <- ggraph::ggraph(net, layout = 'star') +
    ggraph::geom_edge_arc(color = "green", width = 0.7) +
    ggraph::geom_node_point(size = 5, color = "gray50")
p2
```

and...

```{r dynamic2}
p3 <- ggraph::ggraph(net, layout = 'linear') +
    ggraph::geom_edge_arc(color = "blue", width = 0.7) +
    ggraph::geom_node_point(size = 5, color = "gray50")
p3
```


At this point we need to change the header of the code box to  
_r name-of-your-choice, fig.show='animate', ffmpeg.format='gif', dev='jpeg_
and print by calling the plots by name.

-- you can't see it but the following box contains this header

```{r gapgif, fig.show = 'animate', ffmpeg.format = 'gif', dev = 'jpeg'}

# Display dynamic plots

p1
p2
p3
```


If you are seeing the plots moving, we made it!
Hard to tell from this dynamic viz, but it's the same network!

At this point, we might want to save the plot as a gif file, so it
could be embedded to documents, or shared with colleagues 
and friends (Ok... nerd friends!)

We can save the dynamic viz into a gif in several ways. 
I suggest one relatively easy way to do it here. 

* Download the software _ImageMagick_ here 
https://imagemagick.org/script/download.php

* call the library `animation`

```{r dynamic3}
library(animation)
```

* check the pathway of _ImageMagick_
```{r dynamic4}
animation::ani.options("convert")
```

* If your R Studio can't find it automatically, you should find it manually,
place it in a convenient folder and feed a pathway to the following line of code. 
For instance, this is mine, running on Linux (aka each system has a different
way to specify pathways. Make sure you use the right one for your system). 

```{r dynamic5}
animation::ani.options(convert = "home/claudia/Documents/ImageMagick-6.8.8-Q16/convert.exe")
```

When you are done with setting up, we can make a gif! The `saveGIF` function
handles expressions (check the help page). We need to find a way to give it what
it wants. 

* Let's save our plots in a list 
* Write a for loop that goes through the list
* Place the for loop as an argument in the `saveGIF` function

```{r dynamic6}
plots <- base::list(p1, p2, p3)

# for loop to be inserted as the function argument
# for (i in 1:length(plots)){
#   base::print(plots[i])
#   }

animation::saveGIF(
  for (i in 1:length(plots)){
    base::print(plots[i])
  }
  , movie.name = "attempt.gif")
```
You can't see it now, but this code saves the gif in your selected folder. 


### Interactive Networks

Another cool thing we can make is interactive networks. An interactive 
network allows the reader to zoom in and out, move the layout, 
focus on subgraphs, and display extra information 
when pointing a node with the cursor. 

Let's give it a go with the `networkD3` library after making a data frame 
with the kind of data that it likes!

* The nodes IDs have to be numeric, and they also have to start from 0:
An easy way to get there is to transform our character IDs into a factor 
variable, transform it to numeric, and make sure it starts from zero 
by subtracting 1.

* Then we can customize the node names with an attribute such as media, 
provided it is a factor variable


```{r dynamic7, message = FALSE, warning = FALSE}
library(networkD3)
```

```{r dynamic8}
# make node IDs numeric
links.d3 <- base::data.frame(from = as.numeric(factor(links$from))-1, 
                             to = as.numeric(factor(links$to))-1 )

# Uses the name of the media as node ID used as a factor variable.  
# We create the idn variable so we do not overscribe the original variable 
# a factor
nodes.d3 <- base::cbind(idn = factor(nodes$media, levels = nodes$media), nodes) 

networkD3::forceNetwork(Links = links.d3, Nodes = nodes.d3, Source = "from", Target = "to",
                        NodeID = "idn", Group = "type.label",linkWidth = 1,
                        linkColour = "#afafaf", fontSize = 12, zoom = T, legend = T,
                        Nodesize = 6, opacity = 0.8, charge = -300, 
                        width = 600, height = 400)
```

### Let's make a movie!

I'm pretty sure we are not going to win an Oscar, but it's still pretty cool to 
make a movie with network viz!

The `ndtv` package can help us to do it! 

`ndtv` is part of the `statnet` family and it takes class `network` objects. 
Hence we will use one of the `statnet`'s toy data set to try out.
The Florentine Marriage is a classic of Social Network Analysis. It describes 
the way the Medici family got to rule Florence about 500 years ago. 
They strategically married with the most powerful other families 
in order to make alliances and rule the city. 

```{r dynamic9, message = FALSE, warning = FALSE}

library(ndtv)
library(statnet)

```

Temporal Data

```{r dynamic10}
utils::data(short.stergm.sim)
base::class(short.stergm.sim)
utils::head(as.data.frame(short.stergm.sim))
```

This is a temporal edge list:

* An edge goes from a node with ID in the tail column 
* to a node with ID in the head column. 

* Edges exist from time point onset to time point terminus. 

* Each of those periods is recorded on a separate row in the data frame above.

IMPORTANT: The idea of onset and terminus censoring refers to start and 
end points enforced by the beginning and end of network observation rather
than by actual tie formation/dissolution


Here is the Florentine Marriage movie!

```{r dynamic11, results = 'asis'}

ndtv::render.d3movie(short.stergm.sim,
                     displaylabels=TRUE,
                     output.mode = 'inline' #,
                     # filename = "Florentine.html if yo run this argument it opens
                     # the movie on browser and it saves the movie on you set folder.
)
```

Take into account the box header needs to be set to
_r name-of-your-choice,results = 'asis'_.

Very complex visualizations can be displayed with the `ndtv` and we might use it 
again later in the semester. However, we need to grasp a better understanding of
temporal network data sets in order to properly master this sort of visualizations.

### A tiny bit jump into temporal data

DIY is the best way to learn something. Let's make temporal data from scratch!

```{r dynamic12, message = FALSE, warning = FALSE}
t0<-network::as.network(matrix(c(0,1,0,
                        0,0,0,
                        1,0,0),ncol = 3,byrow = TRUE))

t1<-network::as.network(matrix(c(0,1,0,
                        0,1,0,
                        0,0,0),ncol = 3,byrow = TRUE))

t2<-network::as.network(matrix(c(0,0,0,
                        0,1,0,
                        0,1,0),ncol = 3,byrow = TRUE))
# convert a list of networks into networkDynamic object
tnet<-networkDynamic::networkDynamic(network.list = list(t0,t1,t2))

```


Prepare the pop-corn!
```{r dynamic13,results = 'asis'}
ndtv::render.d3movie(tnet,
                     output.mode = 'inline')
```

It worked! 

For this week this is enough! We can make more complicated viz later on in 
the semester!




### Conclusive Remarks

Needless to say, making good visualizations takes plenty of time, and to a
certain extent, it is an art! Try to express your creativity and to have 
fun when you do it! 
