---
title: "lab05_DataManipulation"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: true
    allow_skip: true
    # toc: true
    # toc_depth: 2
    theme: readable
runtime: shiny_prerendered
---
```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome

Hi there, welcome to the tutorial about data manipulation. 
As you might have noticed at this point, data doesn't come 
ready made to be analyzed with you favorite package. 
There are 

* Several packages that require different data format
* Different functions in the same package that ask you 
to manipulate the data in different ways in order to make 
it readable for the package. 

For this reason, in this tutorial we will cover basic data
manipulation in `r` and different ways to load your data in 
the `igraph` package. Many of the formats we cover work
also for the `network` package. 

## Data Frames in r

When you collect data, or when you get data from some data base
you usually receive a spreadsheet. The simplest way to handle 
that data and make it in the shape you need, consists of loading 
it as a data frame. 

On of the most used format is `csv`. 

`r` base allows you to import that format directly with the 
function `read.csv`. 

For instance ` read.csv("some_data.csv", header = TRUE, sep = ",")`

It is very often a good idea to specify that the first row 
should be read like a header, rather than an observation setting` header = TRUE`. 
It is also important to specify how the entries are separated. 
One of the most common is the comma, as specified by `sep = ","` in the example
above.

Explore the other options that you can specify with `?read.csv`. 
Your goal is to read your spreadsheet in the most functional way. 
Correctly importing data usually saves a LARGE amount of time. 

Now, just to familiarize with the r language, let's work
with some examples of data manipulation.

Let's create a data frames in `r`

```{r DfGen, echo = TRUE, include = TRUE, exercise = TRUE}
(df <- data.frame(col_one = LETTERS[1:10], col_two = rep(1:2, 5)))
```

Let's pay attention to a few things:

* You can name your column directly inside the `data.frame`
function, as I just did here with the labels `col_one` 
and `col_two`

* `LETTERS` is a vector of capital letters already stored in
the `r` environment. Using squared brackets I accessed the
first ten elements of the vector

* the function `rep` replicates the values before the comma, 
as many time as indicated by the value after the comma. If 
it was`1:3`the function would have replicated `1, 2, 3` 
for 5 times. 
* the round parenthesis surrounding the all function tells
`r` to print the stored data frame too.

How would you have created the df data frame not using 
any of the shortcuts I showed you above?


```{r grade_df1, exercise = TRUE}


```

```{r grade_df1-solution}
df <- data.frame(c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"), c(1, 2, 1, 2, 1, 2, 1, 2, 1, 2))
colnames(df) <- c("col_one", "col_two")
df
```

```{r grade_df1-check}
gradethis::grade_code(correct = "Well Done!")
```


Sometimes you have a data frame and you need to bind a 
vector as a column. In this case you use the function
`cbind`.

```{r anotherBind, exercise = TRUE}

df <- data.frame(col_one = LETTERS[1:10], col_two = rep(1:2, 5))
col_three <- rep(0,10)

(df <- cbind(df, col_three))


```

It can also happen that you need to bind rows to your 
data frame. You can do it with the function `rbind`

```{r grade_rbind, exercise = TRUE}

df_A <- data.frame(col_one = rep('A',5), col_two = rep(1, 5))
df_B <- data.frame(col_one = rep('B',5), col_two = rep(2, 5))

# how do you put them together and print them in one go?

```



```{r grade_rbind-solution}

df_A <- data.frame(col_one = rep('A',5), col_two = rep(1, 5))
df_B <- data.frame(col_one = rep('B',5), col_two = rep(2, 5))

(df <- rbind(df_A, df_B))

```


```{r grade_rbind-check}
gradethis::grade_code(correct = "That's right!")
```



But take into account that `rbind` works only if the columns 
have the same name in the two data frames you are binding
together. If they don't, you need to rename the columns first.
You can do it with `colnames`. If you need to name rows, guess 
what? You have to use the function `rownames`
```{r colnames, echo = TRUE, include = TRUE, exercise = TRUE}

df <- data.frame(c('A', 'B'), c(1, 1))

colnames(df) <- c('one', 'two')
rownames(df) <- c('one', 'two')

df

```


If you didn't get it right in the exercise above 
where you had to create a data frame with no shortcuts
it's probably because you didn't use `colnames` to 
name the columns. Go back and try again now that 
you know how to do it!


One more important thing is accessing data frames. 
You can do it in two ways:

* Dollar sign `$`
* Squared Brackets `[x, y]`

If you want to access column two you can either
```{r accessCol, echo = TRUE, include = TRUE, exercise = TRUE}
df <- data.frame(col_one = LETTERS[1:10], col_two = rep(1:2, 5))


df$col_two

# or
df[ , 2]


```

these two options print out the same outcome. 

You can also print out a single or a few rows or
modify them

```{r accessRow, echo = TRUE, include = TRUE, exercise = TRUE}

df <- data.frame(col_one = LETTERS[1:10], col_two = rep(1:2, 5))

df[3, ] # prints out row three

df[3:7, ] # prints out rows from 3 to 7

df[3:7, 2] <- c(rep(0,5)) # replaces the values from 3 to 7 in column 2
df

```

When you analyze networks, you often need to deal with 
matrices rather than data frames. You can easily 
turn your data frame into a matrix and back into 
a data frame.

```{r accessRowTwo, echo = TRUE, include = TRUE, exercise = TRUE}
df <- data.frame(one = c('A', 'B'), two = c(1, 1))
class(df)
class(df$two)

df <- as.matrix(df)

class(df)

df <- as.data.frame(df)

class(df$two)

```

Even if df is a data frame again, do you see 
any difference in the class of column two?

Matrices can store only one type of variables.
Since the original df has both character and numeric, `r` 
turned the numeric into character. 
It's obvious that it could not turn the character into
numeric. However, some times 
`r` turns it into factors.
Be careful when you transform the class of your
objects, since this could be the cause of bugs
in your code!

Enough with data frames, let's move to something 
juicier!

## Making networks

When we have new data that no one prepared for us
it is essential to make it in the right shape to 
be loaded into the package that we are using. 

Usually row network data collected with
a survey looks like this


```{r load_FakeNetActors, include = FALSE}
FakeNetActors <- SNA4DS:::FakeNetActors
```


```{r FakeNetActors_explore, exercise = TRUE, exercise.setup = "load_FakeNetActors"}
# the data is loadead in another box you cannot see. If it doesn't load re-install SNA4DS
FakeNetActors

```

This is a completely fake data set that I created for this class, using 
known names to make it a little easier to discuss. 
Would you rather discuss node A behavior's or George Clooney? 
I'll take Clooney, even if this behavior is a fake one. 

So let's assume that this network tells us about actor's friendship and 
that I collected this data when I went to Holliwood interviewing each 
of them personally. 

When you collect data using a survey, either in person or online, you usually 
present the data in this format:

In the first column there is always information on 
the survey respondents. 

In this example the second column reports a node attribute, respondent's sex.
It's common to find information about node attribute next to the respondent 
column. However, sometimes is after the alters. 

What are the alters? The friends. The survey non-respondents that are
still considered in the sample.

How to select your sample, and how to decide whether to keep or not to 
keep an alter in your study requires a very important and long 
discussion. However, here we are just learning how to manipulate 
data and we assume that these alters make perfect sense and that knowing 
about two best friends of the respondent is a fantastic research strategy. 

If we had only one alter this would be already an edge list and you 
should just change the order of the columns to import it into 
`igraph`. When you have more than one alter, you have to manipulate
the data. The idea is that each alter column has to match with the 
respondent column. If a data set is simple and small like that you 
could do it manually. BUT, we are programming nerds, and we don't 
do manual work! Plus, when you have 4000 nodes and 10 alters you 
don't want to do this manually. 
In this package I added a function that does this for you: 
`SNA4DS::makeEdgeList`



```{r FakeNetActors_edgelist, exercise = TRUE, exercise.setup = "load_FakeNetActors"}
(edgelist <- SNA4DS::makeEdgeList(names = FakeNetActors[, c('respondent', 'friend_one' , 'friend_two' )]))

```




```{r load_FakeNetActorsedgelist, include = FALSE}
edgelist <- SNA4DS:::FakeNetActorsedgelist

```

Here is your ready made edgelist. The function can also handle 
one edge attribute at a time, that would pop as a third column
in your output. Don't worry, you'll get a chance to try that out!
To understand how to use the function, as usual, 
type `?SNA4DS::makeEdgeList`

You also need a nodelist. If your alters are also respondents, 
this step is not necessary. However, this is very rarely the case.
To get a nodelist you can use `SNA4DS::makeNodeList`.


```{r load_FakeNetActornodelist, include = FALSE}

nodelist <- SNA4DS:::FakeNetActorsnodelist
```

```{r FakeNetActors_nodelist, exercise = TRUE, exercise.setup = "load_FakeNetActors"}
(nodelist <- SNA4DS::makeNodeList(names = FakeNetActors[, c('respondent', 'friend_one' , 'friend_two' )], attribute = FakeNetActors$sex))

```

In this case, there is an attribute and the function is processing it. 
As you can see we don't have any info on the sex of the alters, since 
they did not answer the survey. Hence, the function returns missing data.

At this point we are ready to load our data into igraph with the function
`igraph::graph_from_data_frame`. 


```{r grade_FakeNetActors_makegraph, exercise = TRUE, exercise.setup = "load_FakeNetActors"}

ActorsNet <- igraph::graph_from_data_frame(d = edgelist, vertices = nodelist, directed = T)
plot(ActorsNet)
  
```





In case you have more than one edge or node attribute to process, 
you can run the two functions as many times as the number of attributes to process
and then `cbind` the columns of the data frames before loading them
into `igraph`.

## Destroy and Build Again

There are several other ways to load data into `igraph`. 
Very often you get network data in weird shapes and you need
to find the easiest and quickest way to make into a network.
Sometimes you also need to extract bits from an object of class
network too. 

Let's take our `ActorsNet` and decompose it into pieces, 
and load it into igraph again in different ways so we can 
learn how to handle this kind of problems.

```{r extractMat, exercise = TRUE, exercise.setup = "load_FakeNetActors"}
ActorsNet <- igraph::graph_from_data_frame(d = edgelist, vertices = nodelist, directed = T)

(ActorAdjMat <- igraph::as_adjacency_matrix(ActorsNet))

AdjMatGraph <- igraph::graph_from_adjacency_matrix(ActorAdjMat)

# extract attribute

sex <- igraph::V(ActorsNet)$attribute

# add attribute to the graph made with the adj mat

igraph::V(AdjMatGraph)$sex <- sex

AdjMatGraph
```
Ok, what happened there? we took the graph we made with the 
FakeNetActors and we extracted the adjacency matrix. 
We made it into a graph again, but this time using 
`igraph::graph_from_adjacency_matrix`. However, the original
graph also had one attribute. We extracted it from the original 
graph and imported into the new one changing the name. 

Even if this might seem a tedious game, when you have to make 
transformation on complex graph this sort of transformations 
are your bread and butter.

If I tell you that there is a way to do the same procedure 
extracting the edgelist and loading the graph as an edge list, 
and if I also tell you that you can use these two functions
`igraph::get.edgelist`
`igraph::graph_from_edgelist`
Would you be able to do it?


```{r extractEdgeList, exercise = TRUE, exercise.setup = "load_FakeNetActors"}
ActorsNet <- igraph::graph_from_data_frame(d = edgelist, vertices = nodelist, directed = T)

ActorEdL <- 

EdgeListGraph <- 


```

```{r extractEdgeList-solution}
ActorsNet <- igraph::graph_from_data_frame(d = edgelist, vertices = nodelist, directed = T)

ActorEdL <- igraph::get.edgelist(ActorsNet)

EdgeListGraph <- igraph::graph_from_edgelist(ActorEdL)

```



```{r extractEdgeList-check}
gradethis::grade_code(correct = "Well done you!")
```

That worked very well! Well done, BUT... 
there is a but. Take into account that here it worked well
since there are no isolates. If there were isolates, those
wouldn't be in the edgelist, since they have no edges and 
you would get missing nodes in your network. 

That doesn't happen with the adjacency matrix, but if you have about
100k nodes, it's computationally heavy. Using the data frame
as we did at first addresses both problems. Hence, whenever it
is possible, go for that!


## Apollo 13!

Now we are going to work with some data from the Apollo 13
mission. They went to the Moon! How nerdy cool is that?!

The data is freely available online in `excel` format. 
`r` doesn't speak `excel`, so we need to use a library to
allow `r` to understand our Apollo astronauts. 
Among many available packages that do so, I suggest `openxlsx`.


After installing the library, opening it and checking all 
the available options with the help function, you can use 
a line of code as such to load the data:

`Apollo13 <- read.xlsx("Apollo13_data.xlsx")`

However, for this tutorial, the data is already loaded
for you into `SNA4DS` and you can also open it in your 
`rstudio` console later directly from the package. 

You are welcome :)

```{r load_apollo13, include = FALSE}
Apollo13 <- SNA4DS:::Apollo13
```


Since the data is already there let's take a look and
observe its structure using the `r` inbuilt function `head`


```{r apollo13_dataOverview, exercise = TRUE, exercise.setup = "load_apollo13"}

head(Apollo13)

```
In this data frame we have five columns. We already know by heart columns
three and five. They constitute our edgelist, providing the sender and 
the receiver... wait, sender and receiver of what?! ah a message. So in 
this data set nodes are people involved in the Apollo13 mission, and edges
are messages they sent to each other. Since these messages are MANY, we also 
have a column that tells us when these messages where sent.

By checking the length of the data frame, we can determine how many messages
(edges) we are dealing with.

```{r apollo13_nrow, exercise = TRUE, exercise.setup = "load_apollo13"}

nrow(Apollo13)

```
 
Waw! That's quite some messages! To start with, it's a good 
idea to subset some of them and work with those only. 

Obviously you normally subset according to a criteria 
that it is relevant to your research question. However, 
this is just a tutorial on data manipulation! We don't have
a research question!

Let's subset a sample of this data set that contains the word
Roger! --I'm gelous, the word Claudia is not there ;)

```{r apollo13_Roger, exercise = TRUE, exercise.setup = "load_apollo13"}

roger <- Apollo13[grep('Roger', Apollo13$Message), ]
head(roger)
```

With this line of code we subsetted the roger-Apollo13 data set. 
Just to make sure, Roger stands for 'received in radio traffic', 
not for Leenders :)

What happened in that line of code? we are asking `r` to subset only
the rows that contain the word `Roger` in the Apollo13 data set.
`Apollo13[here, ]` is selecting rows. To specify we want those containing 
`Roger` we use the `grep` function, that fishes a pattern for us. 

Why don't you try to play with the above code and fish some other pattern?
Try something simple such as 'the'. 

You can also decide to subset the Apollo13 data frame according to 
smaller time windows. Why don't you try to subset only the time stamps 
that start with `2.32`?


```{r apollo13_time, exercise = TRUE, exercise.setup = "load_apollo13"}

TimeWindow <- 

```


```{r apollo13_time-solution}

TimeWindow <- Apollo13[grep('2.32', Apollo13$Time), ]

```



```{r apollo13_time-check}
gradethis::grade_code(correct = "You did it! Cool :)")
```


Time stamped data is really cool when it comes to network models.
There are many different things you can do with that, among which
observing the evolution of a network over time.

We could split the data set into time units and set the number of
messages between two nodes as the weight. Then it would be nice to 
plot al these time stamps as networks next to each other and observe 
the evolution of the Apollo13 conversations over time!!

We could, but for now, this is to difficult. But you'll get there, 
because remember 'Failure is not an option'! ... And that's the motto 
of the Apollo missions! 


## The tidyverse partition

If you run a community detection algorithm on the network
depicting `r` users you will find a major split. 
On the one side there are people that use `r`base code, on the 
other there are the `tidyverse` folks and the polarization is huge.

`tidyverse` is a project aimed at changing the way people code in `r`.
More info are available [here](https://www.tidyverse.org/). Some 
people say it is the equivalent to `Panda` in `Python`. It's just a 
matter of personal taste at the end. But let's have a look at how 
the `tidyverse` planet revolves around networks using the `media_net`data again!

Let's load the data
```{r load_media_net, include = TRUE, echo = TRUE}
media_net <- SNA4DS:::media_net

```


First we need to transform the class `igraph` into `tbl_graph`. 
Then we can use `tidyverse` to subset parts of the data set.

```{r media_net_tidyverse, exercise = TRUE, exercise.setup = "load_media_net"}
library(tidyverse)

media<-tidygraph::as_tbl_graph(media_net)
# shows info on edges
media %>%  tidygraph::activate(edges) %>%  tidygraph::as_tibble()
# there are two types of edges, here we filter for one of them
# store the output
# print it
(media_edges <- media %>% tidygraph::activate(edges) %>% tidygraph::filter(type=="hyperlink"))
```


The function `activate` is part of `tidyverse`, which is
a collection of packages. But `tidyverse` does not export 
functions per se, and when you use a function, you need to
find out which is the package that contains it. In this case
it's `tidygraph`, as much as many other functions that you
use dealing with networks. 

As a `tbl_graph` can be considered as a collection of two linked 
tables it is necessary to specify which table is referenced during 
manipulations. The `activate` verb does just that and affects 
all subsequent manipulations until a new table is activated. 
It is a simple query function to get the currently active context. 

We used `activate` to access information about edges and edges
attribute (type). This can be useful both to quickly access info
you might need to see or display, and for plotting. We can talk 
about plotting in a minute. First, can you access information 
about nodes first and filter by newspaper afterwards?

```{r tidyverse_nodes, exercise = TRUE, exercise.setup = "load_media_net"}

library(tidyverse)

media <- tidygraph::as_tbl_graph(media_net) 

# don't forget to print the output

```


```{r tidyverse_nodes-solution}

library(tidyverse)

media <- tidygraph::as_tbl_graph(media_net)

media %>%  tidygraph::activate(nodes) %>%  tidygraph::as_tibble()

# don't forget to print the output

(media_nodes <- media %>% tidygraph::activate(nodes) %>% tidygraph::filter(media.type=="newspaper"))

```



```{r tidyverse_nodes-check}
gradethis::grade_code(correct = "Super!")
```


`tidyverse` is particularly efficient to make pretty plots. Let's try it out!

```{r pretty_plot, exercise = TRUE, exercise.setup = "load_media_net"}
library(ggraph)
library(tidyverse)
media <- tidygraph::as_tbl_graph(media_net)

media_m<-media %>% tidygraph::activate(nodes) %>%  dplyr::mutate(graph="Mention")

media_e <- media %>% tidygraph::activate(edges) %>% tidygraph::filter(type=="hyperlink")
media_e1<-media_e %>% tidygraph::activate(nodes)%>%  dplyr::mutate(graph="Hyperlink")

ggraph(tidygraph::bind_graphs(media_m,media_e1), layout = 'nicely') + 
    ggraph::geom_edge_link(arrow = arrow(length = unit(2, 'mm')), 
                   end_cap = circle(3, 'mm')) + 
    ggraph::geom_node_point(size = 3) +  
    ggraph::geom_node_text(aes(label = media, color = type.label), size=3, repel = T) +
    ggraph::facet_nodes(~graph) + 
    ggraph::theme_graph(foreground = 'steelblue')
```

In this chunk of code we are doing LOTS of things at 
the same time!

1) We again set the class of the graph to `tbl_graph`.
2) We activated the chunk of code related to nodes and we create a new column 
using the function `dplyr::mutate` to create a label for each of the nodes,
that will be used for the plot later
3) then we activated edges of the 'hyperlink' type only
4) Plus, in reference to the activated edges, we also activated the info 
related to nodes, while creating a new column that works as a label. 
5) We plotted the network in the usual way, specifying two new options:
* the fact that the two subset we created are going to be plotted together
`tidygraph::bind_graphs`.
* that there are two facets in the plot, each one showing one of the two 
chunks we prepared

The facet plots can be super useful plotting temporal data with time stamps!



### Enough for today!

See you for the next tutorial! :)








