---
title: "Lab 03 - Measures I"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: true
    allow_skip: true
    # toc: true
    # toc_depth: 2
    theme: readable
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction
Welcome to the third tutorial in the series of __Social Network Analysis for Data Science__!
This week is devoted to an introduction to measuring networks in r. 
Another tutorial, Measure II, will follow next week.

In this tutorial, you will familiarize with network data and, 
try to code some of the topics we covered in the class.

### Let's get started


```{r testquiz}
question("Why do we measure networks?",
    answer("to learn and apply data science"),
    answer("because it's important to make detailed reports of the object of your study"),
    answer("to better understand what kind of relationships they represent 
and what does this imply for the understanding of the social reality.", correct = TRUE)
  )
```


## Network structures
In class, we covered some standard network structures. 
Let's explore them together.

`igraph` allows you to generate networks with regular structures.

In the following example, we use the function `igraph::make_full_graph` 
to generate a fully connected graph with 10 nodes. 

### Full Graph
```{r full_graph, echo = TRUE, include = TRUE, exercise = TRUE}
library(igraph)
full <- igraph::make_full_graph(10, directed = FALSE, loops = FALSE)

plot(full)

```


The previous chunk of code generated and plotted a fully connected graph with 
10 nodes. Try to substitute 10 with 100 and 3, and observe what happens if you 
reduce and increase the number of nodes. 

Independently of the number of nodes you choose, it is going 
to be fully connected. 

The graph in the previous chunk is undirected. However, you 
can also try to generate and plot a directed graph, by setting 
the parameter `directed` to `TRUE`. 


### Star
In the following chunk, we generate another standard shape, 
the star.  

Please note that I'm skipping a passage and putting the star
function directly into the plot function. This implies that 
my star is not stored as a variable in my `r` environment.
In this case, it will print, that it's what I need now,
but it won't be saved for future work. 


```{r star, echo = TRUE, include = TRUE, exercise = TRUE}
plot(igraph::make_star(10, mode = "out"))

```


This time we generated a directed graph. In this case too, 
you can have different kinds of ties. But the way to do it 
is different from the previous function even if they both 
belong to `igraph`.
There is no common standard due to the participation of 
different developers that contributed to the package. 
This means that you need to check the help file every time. 
In this case, you use `?igraph::make_star`. 
There are three more possible arguments after "out" to be 
specified with the parameter `mode`: "in",  "mutual", "undirected".
Try them and see what happens before moving on to the next regular
structure!


### Ring

Our next shape is the Ring

```{r ring, echo = TRUE, include = TRUE, exercise = TRUE}
plot(igraph::make_ring(10, directed = TRUE, mutual = TRUE))
```

This is also directed. To make it undirected you need to set two 
parameters to `FALSE`.

### Linear
The next regular shape we plot is a linear one. The linear shape is
just like a ring that is missing one edge. 

That's exactly how we are going to generate it then!


```{r linear, echo = TRUE, include = TRUE, exercise = TRUE}
# removing edge to make it linear
ring <- igraph::make_ring(10)
igraph::V(ring)$name <- letters[1:10] # naming nodes
igraph::E(ring)$name <- LETTERS[1:10] # naming edges
ring <- ring - igraph::edge("e|f") # removing the edge that goes from node e to f
plot(ring)

```
We could also have removed one edge calling it by its name.
Try to add up this further chunk of code and see what happens:
`ring <- ring - edge("A")
plot(ring)`


### Tree

As we have seen in the lecture, a tree is a peculiar network type. 
If you type `?igraph::make_tree` in your console, you find out that
the second parameter you pass on is the number of children. In the
following example, there are two of them.

```{r tree, echo = TRUE, include = TRUE, exercise = TRUE}
plot(igraph::make_tree(10, 2, mode = "out"))
```

While the first parameter is the number of nodes, the
`children` parameters is setting the number of connection of a node. 
In this case, the number of `out` connection. Try also 'in' and
'undirected' and see how the tree changes. 


### Lattice

The lattice is one of the most fascinating shape we will ever 
study in network theory. The following code prints a squared one

```{r lattice, echo = TRUE, include = TRUE, exercise = TRUE}

# squared lattice
plot(igraph::make_lattice(length = 5, dim = 2))

# plot(igraph::make_lattice(c(5, 5, 5)))
```

But try also to comment the first line of code and make the second line run
to observe how a lattice can also be!

### Combine Graphs

These shapes have a theoretical relevance per se, but they
are also helpful if we want to create graphs with specific
combined shapes.
```{r combine, echo = TRUE, include = TRUE, exercise = TRUE}

G1 <- igraph::make_tree(10, 2, mode = "out")

G2 <- igraph::make_star(10, mode = "out")

net <- igraph::union(G1,G2)
# same as above written differently -it might not run in the tutorial
# net <- G1 %u% G2 

plot(net)

```


We can also add up more nodes and links in other ways

```{r combine_literal, echo = TRUE, include = TRUE, exercise = TRUE}
net1 <- igraph::graph_from_literal(D-A:B:F:G, A-C-F-A, B-E-G-B, A-B, F-G,
                  H-F:G, H-I-J)
net2 <- igraph::graph_from_literal(D-A:F:Y, B-A-X-F-H-Z, F-Y)


net <- igraph::union(net1,net2)
# same as above written differently -it might not run in the tutorial
# net <- net1 %u% net2

plot(net)

```



## Watch the Matrix!


Networks (graphs) are matrices, and they speak matrix algebra.
The easiest way to CLEARLY understand network structures is from 
observing a matrix.

Let's generate a 3 x 3 matrix that represents a network with 3 
nodes

```{r mat_create, echo = TRUE, include = TRUE, exercise = TRUE }
mat <- matrix(0, 3, 3)

mat
```




If a network is undirected, the matrix is symmetrical along the 
diagonal since the tie is reciprocal.



```{r mat_sym, echo = TRUE, include = TRUE, exercise = TRUE }
mat <- matrix(0, 3, 3)
mat[2,1] <- 1
mat[1,2] <- 1
mat[3,2] <- 1
mat[2,3] <- 1

mat
isSymmetric(mat)
```
if we import it into `igraph` we can plot it 

```{r mat_sym_plot, echo = TRUE, include = TRUE, exercise = TRUE }

mat <- matrix(0, 3, 3)
mat[2,1] <- 1
mat[1,2] <- 1
mat[3,2] <- 1
mat[2,3] <- 1
g <- igraph::graph_from_adjacency_matrix(mat, mode = 'undirected')
plot(g)
```


In case the matrix is not symmetrical, we should plot it as a direct graph.

```{r mat_on-sym, echo = TRUE, include = TRUE, exercise = TRUE}
# make a not symmetrical matrix

mat <- matrix(0, 3, 3)
mat[1,3] <- 1

g <- igraph::graph_from_adjacency_matrix(mat, mode = 'directed')
plot(g)

```

Note that `ìgraph` doesn't know if your data is symmetrical or not. 
If you set 'undirected' for a non-symmetrical matrix, it will plot it
anyway. You usually get this information from the data collection. 
Otherwise, always check the matrix.

```{r mat_non-sym_check, echo = TRUE, include = TRUE, exercise = TRUE,}

mat <- matrix(0, 3, 3)
mat[1,3] <- 1
isSymmetric(mat)
```

When you are unsure about some network theory, my advice is to
generate a small matrix and make experiments to check how it works!

Ah yes sorry about the title, but in case of misunderstanding,
the Matrix movie is not relevant for this section! :)



## Measuring the Media Network

After pure theoretical network shapes, we move back to our 
beloved social sphere and get back to the media network 
we employed in the previous tutorial already. Even if we 
plotted it in many different ways we still know nothing 
about it, and we cannot fully answer any research question 
concerning this network yet.


```{r load_media_net, include = FALSE}
media_net <- SNA4DS:::media_net

```


```{r media_net, echo = TRUE, include = TRUE}
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)

media_net <- igraph::graph_from_data_frame(d=links, vertices=nodes, directed=T) 
media_net
```


What do we know now? We have 17 nodes and 49 edges. For 
what concerns the nodes we know their name, what type of 
media  they are (Newspaper, TV, Online) and the size of 
the audience. About the edges, we know the type and the 
weight. We will come back to the type in the following weeks. 
For now, let's pretend it's one type only.


### Distance in a Network

When we talk about distance in a network, we are looking for
the number of edges that separate one node from another in 
the same graph provided that these two nodes are connected.
This distance between two nodes is called a path, and its length
is the number of links. The sequence of alternating nodes
and edges in a path is called a walk. If it is possible to get 
from one node to the other, the other is reachable. 

If we use `igraph::distances` we get a table with all the 
distances between each pair of nodes.

```{r media_net_distances, exercise = TRUE, exercise.setup = "load_media_net"}
igraph::distances(media_net, mode = "all")

```

Substituting the argument in the `mode` parameter to "in" or 
"out" we get to observe the distances considering these types
of connections only.


### Shortest Path
The shortest path, or geodesic, between a pair of vertices, is
a path with minimal vertices. The `igraph::shortest_paths`
function shows the shortest path from a given node (in the 
example below s01) and all the other nodes.

```{r media_net_shortPath, exercise = TRUE, exercise.setup = "load_media_net"}

igraph::shortest_paths(media_net, "s01", to = igraph::V(media_net), mode =  "all", output = "vpath")
  

```
Note that: 

1) it provides the distance also from the node itself.
2) the distance can be expressed both with nodes or paths. 

Check the help file `?igraph::shortest_paths` to check what the 
parameters `predecessors` and `inbound_edges` do.



### Diameter

The diameter is the longest shortest path in the networks.

```{r media_net_diameter, exercise = TRUE, exercise.setup = "load_media_net"}
igraph::diameter(media_net)

```


The diameter can be useful in many ways. For instance, if 
you have several snapshots of the same network that evolves 
over time and you want to compare how relationships between 
pair of nodes changed at the structural level.

But other measures can also help us compare networks

### Average path length

The average path length tells us the average distance 
between each pair of nodes. 
```{r media_net_meanDist, exercise = TRUE, exercise.setup = "load_media_net"}
igraph::mean_distance(media_net)
```
A fully connected network has average path length of 1, 
since every single node is connected to any other.

### Density and Sparsity

The density of a graph is the ratio of the number of 
edges and the number of possible edges.
Network density tells you the extent to which a network
is connected. We call sparse a network with very few 
connections and dense, one with many connections.

```{r media_net_density, exercise = TRUE, exercise.setup = "load_media_net"}
igraph::edge_density(media_net, loops=TRUE)  

```

A fully connected graph has density 1. 

The function `igraph::edge_density` does not allow to 
specify whether the network is directed or undirected. 

You have to pass the network in the right way. Our
network is directed, so we did right, but if we want
also to check the undirected density we can do it as 
shown in the following example

```{r media_net_distancesUnd, exercise = TRUE, exercise.setup = "load_media_net"}

igraph::edge_density(igraph::as.undirected(media_net), loops=TRUE) 

```

Anyway, remember that measures per se tell you nothing. They only 
help you if they are used to support a theoretical argument.
If I tell you that I had five ice creams, you cannot understand 
what I'm talking about. 
If I say five ice creams in one hour, you would tell me to slow 
down that it's not healthy. 
If I had five ice creams in the last six months, you could ask me if 
I'm on a diet. The context (theory for us) makes the measure meaningful. 


## Node Degree

Another significant measure that helps in understanding networks is 
the degree of the nodes. 
Node degree is the number of edges connected to the node. To measure 
the degree of each node, we use the `degree` function from the `igraph`
package. 

```{r media_net_degreeDir, exercise = TRUE, exercise.setup = "load_media_net"}

# out degree
igraph::degree(media_net, v = igraph::V(media_net), mode = "out", loops = TRUE, normalized = FALSE)
# in degree
igraph::degree(media_net, v = igraph::V(media_net), mode = "in", loops = TRUE, normalized = FALSE)
# total degree
igraph::degree(media_net, v = igraph::V(media_net), mode = "total", loops = TRUE, normalized = FALSE)



```



Since our network is directed, we can specify three options for 
the estimation of the degree. Each line in the output corresponds 
respectively to ''out'', ''in'' and ''total'' degree. 
Again, out-degree is the numbers of ties coming from a 
specific node, in degree, is the number of ties received, 
and the total is the sum of the two.

We can calculate the undirected version too (sometimes 
it helps!)


```{r media_net_degreeUnd, exercise = TRUE, exercise.setup = "load_media_net"}
Umedia_net <- igraph::as.undirected(media_net)

igraph::degree(Umedia_net, v = igraph::V(Umedia_net), mode = "all", loops = TRUE, normalized = FALSE)

```



Pay attention to the fact that undirected degree and total
degree have different results! 
If you consider the network as directed, some ties are counted 
twice if the relationship is not mutual. This condition is not 
valid for undirected, where there can be only one tie between two
nodes.

Remember, if you write the code the way I did now, you are 
just printing the node degree in your console. If you need
to work with that, you need to save it into a variable.


## Weighted Networks

Sometimes we need to describe relationships that are more 
articulated than a simple connection. One situation that we
often encounter in life is the need to understand the intensity
of a relationship. If someone texts you one time only, you 
can hardly call him a friend. If this person texts you three 
times per week, you can claim you are closed, but not as much 
as if they text you three times per hour. In network terms, we 
express this with an edge attribute called weight. 

In our media data set, the weight expresses how often these
media mention or provide a hyperlink connecting to 
another media (type). For now, let's assume that mentioning 
and linking is the same, and let's just check how often
they do so. 



```{r media_net_weight, exercise = TRUE, exercise.setup = "load_media_net"}
library(ggraph)
ggraph::ggraph(media_net, layout = 'stress') + 
  ggraph::geom_edge_arc(aes(width = weight), color = 'grey', strength = 0) + 
  ggraph::geom_node_point(size = 4) + 
  ggraph::geom_node_text(aes(label = media), size = 4, color = "black", repel = T) +
  ggplot2::ggtitle("Media Weighted Plot") +
  ggplot2::theme_void() 
```



From the plot, we can immediately see that some relationships are
more intense than others. A simple visualization can readily help 
us understand the situation and formulate hypotheses on why some 
relationships are more intense than others. 

You need to think in causal terms. Are more intense relationships
the cause or the consequence of some other event? Are they the 
independent or the dependent variable in our causal model? 


Weight is a very crucial and versatile network attribute that
allows several sophisticated network manipulations and it's 
widely employed in network models. 

Weight can be expressed both as an edge attribute and as an
adjacency matrix that rather than being filled with 0s and 1s,
it is filled with the corresponding weights. 

### Strength

When we deal with weighted networks, the degree provides only 
partial information since this measure does not account for 
the weight of the edges, but only for the raw number. 

To address this problem, we can use a measure of strength that
balances degree and weight. 



```{r media_net_strength, exercise = TRUE, exercise.setup = "load_media_net"}
igraph::strength(media_net) 
```


According to the kind of analysis your are doing you should
decide whether it is more helpful to use degree or strength.


## Network Components

Real-life networks are composed of regular shapes that can be 
subsetted and analyzed more closely to make sense of their structure. 
These parts are called components, or subgraphs.



### Subgraphs
We can eliminate some nodes from the network in order to 
focus on the remaining ones that we selected according to
some criteria. The function `igraph::induced_subgraph`, keeps
only the nodes mentioned as an argument.



```{r media_net_subg, exercise = TRUE, exercise.setup = "load_media_net"}

sub <- igraph::induced_subgraph(media_net, c('s01','s02'))
# sub <- igraph::induced_subgraph(media_net, 1:7)

sub
# plot(sub)

```

All the attributes are unchanged, but every other node disappeared. 
You can also call the nodes by number. 
Try to run the second line of code and comment on the first one instead.
Try to plot them too, to observe the subgraph.

Explore the `igraph ` documentation to find other functions to
create subgraphs.



### Cliques

A clique is a fully connected subset of nodes in an undirected graph.
The function `igraph::cliques` allows you to find the number of cliques
comprised of a certain number of nodes that you specify with the `min`
and `max` parameters. 

```{r media_net_cliques, exercise = TRUE, exercise.setup = "load_media_net" }
igraph::cliques(media_net, min = 4, max = 6) ## all the cliques beween min e max n nodes

# igraph::largest_cliques(media_net)
# 
# igraph::max_cliques(media_net)
```

Try to change the number of nodes and try the other two functions.


## Assortativity

Assortativity represents to what extent nodes in a
network associate with other nodes in the network, 
being of similar sort or being of opposing kind.

One of the most popular measures of assortativity
concerns to the degree of the nodes.


### Assortativity degree
```{r media_net_assortDeg, exercise = TRUE, exercise.setup = "load_media_net"}
# random graph
igraph::assortativity_degree(igraph::sample_gnp(10000, 3/10000))

# preferential attachment
igraph::assortativity_degree(igraph::sample_pa(10000, m=4))

# media network
igraph::assortativity_degree(media_net)
```


A random network has an assortativity degree level closed
to zero; while a real network is more likely to show
some pattern since nodes tends to be connected 
for some reason. Networks characterized by
preferential attachment tend to be dissortative.

Aka, nodes have things in common.

We can also check the assortativity according to
nodes attributes. For instance, according to the 
audience size of each media.

```{r media_net_assort, exercise = TRUE, exercise.setup = "load_media_net"}

igraph::assortativity(media_net, igraph::V(media_net)$audience.size, directed=T)

```

If two nodes are connected and represent people 
that share the same behavior we talk of homophily. 



## Bipartite Networks

Bipartite networks are graphs that express relationships
between two sets of objects, such as people and clubs. 
In contrast, one mode networks only focus on one set of
objects. For example, people only. If we extend the concept to 
N set of objects, we will have N-partite graphs.  

`igraph` allows us to create bipartite data. Let's try out!

```{r bipatite_gen, echo = TRUE, include = TRUE, exercise = TRUE}

g <- igraph::make_full_bipartite_graph(6,3)
g

```

We can also plot it with a special layout for bipartite

```{r bipatite_plot, echo = TRUE, include = TRUE, exercise = TRUE}

g <- igraph::make_full_bipartite_graph(6,3)
plot(g, vertex.label = NA, vertex.size = 7, layout = igraph::layout.bipartite)
```


Try to plot it with another layout for one-mode networks.

We might also want to project our bipartite graph so it represents 
the relationships within one set of object at a time.

```{r bipatite_proj, echo = TRUE, include = TRUE, exercise = TRUE}
g <- igraph::make_full_bipartite_graph(6,3)
proj <- igraph::bipartite_projection(g)
plot(proj[[1]])
# plot(proj[[2]])

```

If it's bipartite, there are two sets, hence two projections.
Try out both. Do you understand the difference?


In the Lab about data manipulation we will explore real
bipartite data.

See you next week!
