---
title: "Lab 03 - Measures I"
output: 
  learnr::tutorial:
    fig_caption: no
    progressive: true
    allow_skip: no
    # toc: true
    # toc_depth: 2
    theme: readable
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(gradethis)
tutorial_options(exercise.checker = gradethis::grade_learnr)
knitr::opts_chunk$set(echo = FALSE)
```


### Introduction
Welcome to the third tutorial in the series of __Social Network Analysis for Data Science__!
This week is devoted to an introduction to measuring networks in r. A more advanced tutorial, 
Measure II, will follow next week.

In this tutorial you will familiarize with network data and, 
try to code some of the topics we covered in the class.

## Let's get started


```{r testquiz}
question("Why do we measure networks?",
    answer("to learn and apply data science"),
    answer("because it's important to make detailed reports of the object of your study"),
    answer("to better understand what kind of relationships they represent 
and what does this imply for the understanding of the social reality.", correct = TRUE)
  )
```


### Network structures

In class we covered some standard network structures. Let's explore them 
together.

`igraph` allows you to generate networks with regular structures.

In the following example we use the function `igraph::make_full_graph` 
to generate a fully connected graph with 10 nodes. 

## Full Graph
```{r full_graph, echo = TRUE, include = TRUE, exercise = TRUE}
library(igraph)
full <- igraph::make_full_graph(10, directed = FALSE, loops = FALSE)

plot(full)

```


The previous chunk of code generated and plotted a fully connected graph 
with 10 nodes. Try to substitute 10 with 100 and 3, and observe 
what happens if you reduce and increase the number of nodes. 

Independently of the number of nodes you choose, it is going 
to be fully connected. 

The graph in the previous chunk is undirected. However, you 
can also try to generate and plot a directed graph, by setting 
the parameter `directed` to `TRUE`. 


## Star
In the following chunk we generate another standard shape, 
the star.  

Please note that I'm skiping a passage here and putting the star
function directly into the plot function. This implies that my 
star is not stored as a variable in my `r` environment.
In this case, it will print, that it's what I need now,
but it won't be saved for future work. 



```{r star, echo = TRUE, include = TRUE, exercise = TRUE}
plot(igraph::make_star(10, mode = "out"))

```


This time we generated a directed graph. In this case too you can 
have different kind o ties. But the way to do it it is different 
from the previous function even if they both belong to `igraph`.
There is no common standard due to the participation of different
developers to the package. This means that you need to check the 
help file every time. In this case you use `?igraph::make_star`. 
There are three more possible options after "out" that you 
specify with the parameter `mode`: "in",  "mutual", "undirected".
Try them and see what happens before moving on to the next regular
structure!


## Ring

Our next shape is the Ring

```{r ring, echo = TRUE, include = TRUE, exercise = TRUE}
plot(print_all(igraph::make_ring(10, directed = TRUE, mutual = TRUE)))
```

This is also directed. To make it undirected you need to set two 
parameters to `FALSE`.

## Linear
The next regular shape we plot is a linear one. The linear shape is
just like a ring that is missing one edge. 

That's exactly how we are going to generate it then!


```{r linear, echo = TRUE, include = TRUE, exercise = TRUE}
# removing edge to make it linear
ring <- igraph::make_ring(10)
igraph::V(ring)$name <- letters[1:10] # naming nodes
igraph::E(ring)$name <- LETTERS[1:10] # naming edges
ring <- ring - edge("e|f") # removing the edge that goes from node e to f
plot(ring)

```
We could also have removed one edge calling it by name.
Try to add up this further chunk of code and see what happens:
`ring <- ring - edge("A")
plot(ring)`


## Tree

As we have seen in the lecture a tree is a very weird network type. 
If you type `?igraph::make_tree` in your console, you find out that
the second parameter you pass on is the number of children. In the
following example there are three of them
```{r tree, echo = TRUE, include = TRUE, exercise = TRUE}
plot(igraph::make_tree(10, 2, mode = "out"))
```

While the first parameter is obviously the number of nodes, the
`children` parameters is setting the number of connection of a node. 
In this case the number of `out` connection. Try also 'in' and
'undirected' and see how the tree changes. 



## Make Lattice

The lattice is one of the most fascinating shape we will ever 
study in network theory. The following code and print a squared one

```{r lattice, echo = TRUE, include = TRUE, exercise = TRUE}

# squared lattice
plot(igraph::make_lattice(length = 5, dim = 2))

# plot(igraph::make_lattice(c(5, 5, 5)))
```

But also comment the first line of code and  make the second line run
to observe how a lattice can also be!

## Combine Graphs

These shapes have a theoretical relevance per se, but they
are also helpful if we want to create graphs with specific
combined shapes.
```{r combine, echo = TRUE, include = TRUE, exercise = TRUE}

G1 <- igraph::make_tree(10, 2, mode = "out")

G2 <- igraph::make_star(10, mode = "out")

net <- G1 %u% G2

plot(net)

```


We can also add up more nodes and links in other ways

```{r combine_literal, echo = TRUE, include = TRUE, exercise = TRUE}
net1 <- igraph::graph_from_literal(D-A:B:F:G, A-C-F-A, B-E-G-B, A-B, F-G,
                  H-F:G, H-I-J)
net2 <- igraph::graph_from_literal(D-A:F:Y, B-A-X-F-H-Z, F-Y)

net <- net1 %u% net2

plot(net)

```

### Watch the Matrix!

Networks (graphs) are matrices and they speak matrix algebra.
The easiest way to CLEARLY understand network structures is from 
observing a matrix.

Let's generate a 3 x 3 matrix that represents a network with 3 
nodes

```{r mat_create, echo = TRUE, include = TRUE, exercise = TRUE }
mat <- matrix(0, 3, 3)

mat
```
If a network is undirected the matrix is symmetrical along the 
diagonal, since the tie is reciprocal

```{r mat_sym,  echo = TRUE, include = TRUE, exercise = TRUE}
mat[2,1] <- 1
mat[1,2] <- 1
mat[3,2] <- 1
mat[2,3] <- 1

mat
isSymmetric(mat)
```
if we import it into `igraph` we can plot it 

```{r mat_sym_plot, echo = TRUE, include = TRUE, exercise = TRUE}
g <- igraph::graph_from_adjacency_matrix(mat, mode = 'undirected')
plot(g)
```


In case the matrix is not symmetrical, 

```{r mat_on-sym, echo = TRUE, include = TRUE, exercise = TRUE}

mat[1,3] <- 1

g <- igraph::graph_from_adjacency_matrix(mat, mode = 'directed')
plot(g)

```

We should plot it as a direct graph. Note that `ìgraph` doesn't know 
if your data is symmetrical or not. If you set 'undirected' for a 
non-symmetrical matrix, it will plot anyway. Normally you get this 
information from the data collection. Otherwise always check the
matrix

```{r mat_non-sym_check, echo = TRUE, include = TRUE, exercise = TRUE}
isSymmetric(mat)
```



Ah yes sorry about the title, but in case of misunderstanding,
the Matrix movie is not relevant for this section! :)



### Media Network

After pure theoretical network shapes, we move back to our 
beloved social sphere and get back to the media network 
we employed in the previous tutorial already. Even if we 
plotted it in many different ways we still know nothing 
about it and we are not able to fully answer any research
question concerning this network yet.


```{r import data}
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)

net <- igraph::graph_from_data_frame(d=links, vertices=nodes, directed=T) 
net
```
What do we know? We have 17 nodes and 49 edgesFor what concerns
the nodes we know their name, what type of media  they are
(Newspaper, TV, Online) and the size of the audience. 
For what concern the edge we know the type and the weight. We
will come back to the type later. For now, let's pretend it's 
one type only.


## Distance in a Network

When we talk about distance in a network we are looking for
the number of edges that separate one node from another in 
the same network provided that these two nodes are connected.
This distance between two nodes is called path and its length
is the number of links. The sequence of alternating nodes
and edges in a path is called walk. If it is possible to get 
from one node to the other, the other is reachable. 

If we use `igraph::distances` we get a table with all the 
distances between each pair of nodes
from ``
```{r distances, echo = TRUE, include = TRUE, exercise = TRUE}
igraph::distances(net, mode = "all")

```

Substituting the argument in the `mode` parameter to "in" or 
"out" we get to observe the distances considering these types
of connections only.


### Shortest Path

The shortest path, or geodesic between two pair of vertices is
a path with the minimal number of vertices. The `igraph::shortest_paths`
function shows the shortest path from a given node (in the example 
below s01) and all the other nodes.

```{r shortest_path, echo = TRUE, include = TRUE, exercise = TRUE}

igraph::shortest_paths(net, "s01", to = V(net), mode =  "all", output = "vpath")
  

```
Note that: 

1) it provides the distance also from the node itself
2) the distance the be expressed both with nodes or paths. 

Check the help file `?igraph::shortest_paths` to check what the 
parameters `predecessors` and `inbound_edges` do.



## Diameter

The diameter is the longest shortest path in the networks.

```{r diameter, echo = TRUE, include = TRUE, exercise = TRUE}
igraph::diameter(net)


# mean_distance(friends)

```

The diameter can be useful in many ways. For instance, if 
you have several 
snapshots of the same network that evolves over time and 
you want to compare how relationships between pair of
nodes changed on the structural level.

But there are other measures that can help us comparing 
networks

## Average path length

The average path length tells us the average distance 
between each pair of nodes. 
```{r mean_dist, echo = TRUE, include = TRUE, exercise = TRUE}
mean_distance(net)
```
A fully connected network has average path length of 1, 
since every single node is connected to any other.

## Density and Sparsity

The density of a graph is the ratio of the number of 
edges and the number of possible edges.
Network density tells you the extent to which a network
is connected. We call sparse a network with very few 
connection and dense one with many connections.
```{r density, echo = TRUE, include = TRUE, exercise = TRUE}
igraph::edge_density(net, loops=TRUE)  

```

A fully connected graph has density 1. 

The function `igraph::edge_density` does not allow to 
specify whether the network is directed or undirected. 

You have to pass the network in the right way. Our
network is directed, so we did right, but if we want
also to check the undirected density we can

```{r density_undirected, echo = TRUE, include = TRUE, exercise = TRUE}

igraph::edge_density(as.undirected(net), loops=TRUE) 

```
Anyway remember that measures per se tell you nothing. They only 
help you if they are used to support a theoretical argument.
If I tell you that I had 5 ice creams you cannot really 
understand what I'm talking about. If I say 5 ice creams 
in one hour you would tell me to slow down that it's not healthy. 
If I had 5 ice creams the all summer you could ask me if 
I'm on a diet. The context (theory for us) makes the measure 
meaningful. 




### Node Degree


```{r}
igraph::degree(friends, v = V(friends), mode = "all",loops = TRUE, normalized = FALSE)

igraph::degree(friends, v = V(friends), mode = "out", loops = TRUE, normalized = FALSE)

igraph::degree(friends, v = V(friends), mode = "in", loops = TRUE, normalized = FALSE)

igraph::degree(friends, v = V(friends), mode = "total", loops = TRUE, normalized = FALSE)

Ufriends <- as.undirected(friends)

degree <- igraph::degree(Ufriends, v = V(Ufriends), mode = "all", loops = TRUE, normalized = FALSE)
# each output gives the same answer

degree_distribution(graph, cumulative = FALSE, ...)

library(xtable)
nodes <- igraph::V(friends)$name <- letters[1:10]

(graph <- data.frame(cbind(nodes, degree)))

```



## Eulerian Path
```{r}

```



### weighted networks

```{r}
igraph::ecount(friends)

igraph::E(friends)$weight <-base::floor(stats::runif(igraph::ecount(friends), min=0, max=20))


E.list <- as.data.frame(igraph::as_edgelist(friends))

E.list <- cbind(E.list, E(friends)$weight)

colnames(E.list) <- c("ego", "alter", "weight")



```

## Strenght
```{r}
strength(friends) ### inserisco???
```




### Network Components

## Subgraphs
```{r}

u2 <- induced_subgraph(friends, 1:7)
plot(friends)
plot(u2)
u3 <- subgraph.edges(u, 1:5, 1:5)



```


## Cliques
```{r}
cliques(friends, min = 2, max = 3) ## all the cliques beween min e max n nodes

largest_cliques(friends)

max_cliques(friends)
```

## Assortativity degree
```{r}
assortativity_degree
```

## Measure Homophiliy

### Bipartite Networks

```{r}
g <- make_bipartite_graph( rep(0:1,length=10), c(1:10))
print(g, v=TRUE)

plot(g)


g <- make_full_bipartite_graph(6,3)
proj <- bipartite_projection(g)
graph.isomorphic(proj[[1]], make_full_graph(10))
graph.isomorphic(proj[[2]], make_full_graph(5))

plot(g)
plot(proj[[1]])
plot(proj[[2]])


```

